---
title: "Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SMARTindicatorsData)
```

# Overview

This package has functions to create and fill templates of ecosystem indicator attributes that will help us determine if they are SMART indicators for a particular purpose. SMART indicators are Specific, Measurable, Achievable, Relevant, and Timely for the priority Council decisions. 

There are two primary workflows to gather indicator SMART attributes. The first automates template filling from the [NEFSC State of the Ecosystem report](https://www.fisheries.noaa.gov/new-england-mid-atlantic/ecosystems/state-ecosystem-reports-northeast-us-shelf) indicators by pulling data from the online SOE dataset [ecodata](https://noaa-edab.github.io/ecodata/index.html), the online Ecosystem Indicator [Catalog](https://noaa-edab.github.io/catalog/), and the online [Technical Documentation](https://noaa-edab.github.io/tech-doc/index.html). The second writes a generic template for a named indicator to a google doc, where a user can fill in the information. Each workflow is explained below. 

Ultimately, indicators from both workflows are combined into a large online book. These workflows and this product were inspired by the [NOAA-EDAB Ecosystem Context for Stock Assessments project](https://github.com/NOAA-EDAB/ECSA). 

This SMART Indicator Reports online book is available as a resource, but each section is summarized into a dataset which can then be searched interactively using the SMART Indicators Dashboard shiny app (currently in development).

# SOE workflow

# Other indicators workflow



