---
title:
    | 
    | SMART Indicator Report
    | {{INDICATOR_NAME}}
#date: "`r format(Sys.time(), '%B %d, %Y')`"
author: ""
always_allow_html: yes
bibliography: smartind.bib
link-citations: true
site: bookdown::bookdown_site
biblio-style: apalike
github-repo: MAFMC/SMARTindicatorsData
header-includes:
  \\usepackage{array}
favicon: favicon.ico
---

```{r, include = FALSE}
## Load packages
# library(tidyverse)
library(ggplot2)
library(tidyr)
library(dplyr)
library(plotly)
library(DT)
library(jpeg)
library(patchwork)
library(ggthemes)
library(here)
library(zoo)
library(gt)
library(kableExtra)
library(ggiraph)

knitr::opts_chunk$set(echo = F, message = F, warning = F, fig.align = 'center', eval.after = 'fig.cap')

## Read in needed data

  ## Select the indicator and check it exists in the list
  ecodatadat <- readRDS(here::here("data-raw/ecodatadat.rds"))

  catalogdat <- readRDS(here::here("data-raw/catalogdat.rds"))

  lookup <- catalogdat |>
    dplyr::filter(Varname == "ecodata name") |>
    dplyr::select(Indicator, Source, Dataset = Value)


## Put all mustaches here
indicator_name <- "{{INDICATOR_NAME}}"
indicator_varname <- "{{INDICATOR_VAR}}"
# cc_name <- "{{CC_NAME}}"
# common_name <- "{{COMMON_NAME}}"
# interactive <-  ifelse(knitr::is_latex_output(), F,T)
# 

## Pull out text and stats for indicator_name

  clean_catdat <- catalogdat |>
    dplyr::filter(stringr::str_detect(Indicator, indicator_name))

  clean_ecodat <- ecodatadat |>
    dplyr::filter(Dataset == lookup$Dataset[lookup$Indicator == indicator_name]) 

  if(!is.null(indicator_varname)){ #no matching needed if doing them all
    clean_ecodat <- clean_ecodat |>
      dplyr::filter(Indicator == indicator_varname)
  }


```


## Indicator category

[Fill by hand, needs decision by Council]

## Indicator name

`r clean_catdat$Value[clean_catdat$Varname==Indicator]`

## Indicator brief description

`r clean_catdat$Value[clean_catdat$Varname==Description]`

## Indicator visualization

```{r}

plotname <- unique(clean_ecodat$Dataset)



```


## Indicator documentation

### Are indicators available for others to use (data downloadable)?

Yes

#### Where can indicators be found?



#### How often are they updated? Are future updates likely?

#### Who is the contact?

`r clean_catdat$Value[clean_catdat$Varname=="Point of contact"]`

### Gather indicator statistics

#### Units

#### Length of time series, start and end date, periodicity

#### Spatial location, scale and extent

#### Management scale: all species, FMP level, species level, can it be aggregated or separated to different scales?

#### Uncertainty metrics


### Are methods clearly documented to obtain source data and calculate indicators?

#### Can the indicator be calculated from current documentation?

#### Is code publicly available? up to date?

#### Have methods changed over time?


### Are indicator underlying source data linked or easy to find?

#### Where are source data stored?

#### How/by whom are source data updated? Are future updates likely?

#### How often are they updated?


## Indicator analysis/testing or history of use

### What decision or advice processes are the indicators currently used in?

### What implications of the indicators are currently listed?

### Do target, limit,  or threshold values already exist for the indicator?

### Have the indicators been tested to ensure they respond proportionally to a change in the underlying process?

### Are the indicators sensitive to a small change in the process, or what is the threshold of change that is detectable?

### Is there a time lag between the process change and the indicator change? How long?


## Comments

### Additional potential links to management in addition to uses listed above


### What additional work would be needed for the Council to use the indicator? 


### What issues are caused if there is a gap or delay in data underlying the indicator




Delete below when done, example from NOAA-EDAB/ESCA https://github.com/NOAA-EDAB/ECSA/blob/master/templates/generic_template1.rmd 

This report provides contextual ecosystem information for {{STOCK_SUBAREA}}{{COMMON_NAME}} (*{{SCI_NAME}}*) on the Northeast U.S. Continental Shelf. Data extractions for spring and fall are confined to the {{COMMON_NAME}} stock area based on respective survey strata sets. The information is intended to span a range of potential factors affecting the productivity and distribution of {{COMMON_NAME}}, including: surface and bottom temperature and salinity, chlorophyll concentrations, indices of habitat, diet composition, and abundance of key zooplankton prey of larval {{COMMON_NAME}}. These factors can be used to qualitatively inform the interpretation of population status and/or quantitatively to improve model responsiveness to ecosystem factors. The range and complexity of ecosystem data makes it unlikely to find the most relevant and comprehensive factor variables with a first evaluation; this process will require an iterative approach of evaluation and feedback. Additional indices can be included to address the needs of the working group.

Link to available [time series data](`r url`) used below.

