---
title:
    | 
    | SMART Indicator Report: {{INDICATOR_NAME}}
    | 
#date: "`r format(Sys.time(), '%B %d, %Y')`"
author: ""
always_allow_html: yes
bibliography: smartind.bib
link-citations: true
site: bookdown::bookdown_site
biblio-style: apalike
github-repo: MAFMC/SMARTindicatorsData
header-includes:
  \\usepackage{array}
favicon: favicon.ico
---

```{r, include = FALSE}
## Load packages
# library(tidyverse)
library(ggplot2)
library(tidyr)
library(dplyr)
library(ecodata)
library(plotly)
library(DT)
library(jpeg)
library(patchwork)
library(ggthemes)
library(here)
library(zoo)
library(gt)
library(kableExtra)
library(ggiraph)

knitr::opts_chunk$set(echo = F, message = F, warning = F, fig.align = 'center', eval.after = 'fig.cap')

## Read in needed data


## Put all mustaches here
indicator_name <- "{{INDICATOR_NAME}}"
indicator_varname <- "{{INDICATOR_VAR}}"

# cc_name <- "{{CC_NAME}}"
# common_name <- "{{COMMON_NAME}}"
# interactive <-  ifelse(knitr::is_latex_output(), F,T)
# 

## Pull out text and stats for indicator_name

  # check for stuff that needs escaping in indicator_name and escape it so detection works
  esc_indicator_name <- stringr::str_escape(indicator_name)



```

{{READMEStart}}

Welcome to this google doc draft SMART indicator report for {{INDICATOR_NAME}}. This document requires hand editing to fill some or all sections. This workflow is inspired by the [Ecosystem Context for Stock Advice](https://github.com/NOAA-EDAB/ECSA/tree/master?tab=readme-ov-file#developing-an-ecsa-project) product workflow. You are looking at an intermediate draft that still needs some work.  

Do:

* Replace the text in [brackets] describing content for the section with the indicator specific content.
* Look at the figures and develop text on how the patterns in the data might be informative to both the assessment and future advice for this stock. 
* Modify the text within sections beginning and ending with text inside double curly-bracket "mustaches". Anything in these sections will be harvested for the final report. 
* Add citations! If you have a citation to add, put it at the end of the section in BibTeX format (.bib; e.g., [here](https://www2.cs.arizona.edu/~collberg/Teaching/07.231/BibTeX/bibtex.html)). 
* Add scientific names, but instead of italicizing, use asterisks before and after, e.g., (*Stenotomus chrysops*).
* Before reading back into an rmd, *all changes must be accepted*

Don't:

* Alter the text within the mustaches! These are highly technical bookmarks that allow us to scrape the text from this google doc into a Rmarkdown (.rmd) file that will be rendered into both .pdf and .html reports. 
* Add new sections. If you want to add a new section, let sgaichas@hydrascientificllc.com know and we can modify the template.rmd at a later date.

{{READMEEnd}}

# Descriptive Section

## Indicator category

{{CategoryStart}}

[Select from: Habitat-Physical; Habitat-Living; Food-Web-Base; Fish; Protected; Fishery; Economic; Social; or Ocean-Use and add between quotes below, then delete this. If multiple categories, add both without punctuation]

{{CategoryEnd}}

`r category <- ""`

`r paste(category)`

## Indicator name

`r paste(indicator_name)`

Includes variable(s): `r paste(indicator_varname)`

## Indicator brief description

{{DescriptionStart}}

[2-3 sentences max plain language overview]

{{DescriptionEnd}}

`r specific.1 <- 1`

## Indicator visualization

{{VizDescriptionStart}}

[What are key trends or patterns we see in the indicator visualization?]

[Add a visualization using a link in markdown format !(caption text)[urllinktofig]]

{{VizDescriptionEnd}}

# SMART Attribute Section

## Indicator documentation

### Are indicators available for others to use (data downloadable)?

{{DownloadableStart}}

[Yes or No, add 1 if Yes and 0 if No to code below]

{{DownloadableEnd}}

`r measurable.1 <- `

#### Where can indicators be found?

{{DataLinkStart}}

[add link to indicator dataset, or state Unknown. Add 1 if link exists and 0 if Unknown to code below]

{{DataLinkEnd}}

`r measurable.2 <- `

#### How often are they updated? Are future updates likely?

{{UpdateFreqStart}}

[Add frequency of updates to date and speculate on continuation of the indicator, or state Unknown.]

{{UpdateFreqEnd}}

#### Who is the contact?

{{POCStart}}

[add contact info for indicator data, or state Unknown.]

{{POCEnd}}

### Gather indicator statistics

#### Units

{{UnitsStart}}

[what are the specific units of each indicator in the dataset, or state Unknown.]

{{UnitsEnd}}

#### Length of time series, start and end date, periodicity

{{TSlengthStart}}

[What is the temporal scale of the indicator? Add specific begin and end years, count total years and count missing years, or state Unknown.]

{{TSlengthEnd}}

#### Spatial location, scale and extent

{{SpatialStart}}

[What is the spatial scale of the indicator? Add specifics if available (lat/long or state boundaries, etc.), or state Unknown.]

{{SpatialEnd}}

#### Management scale: all species, FMP level, species level, can it be aggregated or separated to different scales?

{{MgtScaleStart}}

[Classify by hand, note gridded data if available could be applied to different species ranges, or state Unknown.]

{{MgtScaleEnd}}

#### Uncertainty metrics

{{UncertaintyStart}}

[Yes or no. Are uncertainty metrics included for the indicator (e.g. SD, SE, CI?), if so which ones?]

{{UncertaintyEnd}}

### Are methods clearly documented to obtain source data and calculate indicators?

{{MethodsStart}}

[Yes or no. Are instructions or documentation for calculating the indicator included? Add citation or links]

{{MethodsEnd}}

#### Can the indicator be calculated from current documentation?

{{CanCalcStart}}

[Yes or no. Briefly state what is missing if no.]

{{CanCalcEnd}}

#### Is code publicly available? up to date?

{{CodeAvailStart}}

[Yes or no. Link to indicator calculation code, note most recent update]

{{CodeAvailEnd}}

#### Have methods changed over time?

{{MethodChangeStart}}

[Record any changes in indicator methodology, or state Unknown.]

{{MethodChangeEnd}}

### Are indicator underlying source data linked or easy to find?

{{SourceDataAvailStart}}

[Yes or no. Include link to source data (may be separate from indicator data)]

{{SourceDataAvailEnd}}

#### Where are source data stored?

{{SourceDataStoreStart}}

[Location of indicator source data (may be separate from indicator data), or state Unknown.]

{{SourceDataStoreEnd}}

#### How/by whom are source data updated? Are future updates likely?

{{SourceDataWhoUpdatesStart}}

[Agency and individuals responsible for source data (may be separate from indicator data), or state Unknown.]

{{SourceDataWhoUpdatseEnd}}

#### How often are they updated?

{{SourceDataUpdateFreqStart}}

[Record frequency of updates, or state Unknown.]

{{SourceDataUpdateFreqEnd}}


## Indicator analysis/testing or history of use

### What decision or advice processes are the indicators currently used in?

{{DecisionStart}}

[List details of each process and how the indicator is used. If a current Council process, select the list of management actions/processes using exact keywords]

{{DecisionEnd}}

### What implications of the indicators are currently listed?

{{ImplicationsStart}}

[List the implications of the indicator trends/status/pattern in its current use context]

{{ImplicationsEnd}}

### Do target, limit,  or threshold values already exist for the indicator?

{{RefptsStart}}

[Yes or no. If yes, specifcally what are they?]

{{RefptsEnd}}

### Have the indicators been tested to ensure they respond proportionally to a change in the underlying process?

{{TestedStart}}

[Yes or no. If yes, briefly describe and add citations or links to analyses or previous experience using the indicator]

{{TetstedEnd}}

### Are the indicators sensitive to a small change in the process, or what is the threshold of change that is detectable?

{{SensitiveStart}}

[Yes or no if the answer was yes to testing above. Otherwise, state Unknown.]

{{SensitiveEnd}}

### Is there a time lag between the process change and the indicator change? How long?

{{TimeLagStart}}

[Yes or no if the answer was yes to testing above. Otherwise, state Unknown.]

{{TimeLagEnd}}

# SMART rating

```{r SMART}

indratings <- c("Described" = specific.1, "Units" = specific.2, "Spatial" = specific.3, "Uncertainty" = specific.4, 
                "Methods" = specific.5, "Code" = specific.6,
             "Available" =  measurable.1, "Online" = measurable.2, "Contact" = measurable.3, "SourceDat" = measurable.4, 
             "SourceAvail" = measurable.5, "SourceContact" = measurable.6,
             "Tested" = attainable.1, "Sensitivity" = attainable.2, "TimeLag" = attainable.3,
             "Advice" = relevant.1, "Implications" = relevant.2, "TargThresh" = relevant.3,
             "Frequency" = timebound.1, "Updated" = timebound.2)

ratings <- stack(indratings)

colnames(ratings) <- c("Rating", "Attribute")

ratings$Element <- c(rep("Specific", 6),
                     rep("Measurable", 6),
                     rep("Achievable", 3),
                     rep("Relevant", 3),
                     rep("Timebound", 2))

ratings$Indicator <- unique(clean_catdat$Indicator)
ratings$Category <- category

ratings <- ratings[,c("Category", "Indicator", "Element", "Attribute", "Rating")]

elements <- ratings |>
  dplyr::group_by(Element) |>
  dplyr::summarise(ElementRating = mean(Rating, na.rm=TRUE))

overall <- elements |>
  dplyr::summarise(OverallRating = mean(ElementRating, na.rm=TRUE))

ratingsumm <- ratings |>
  dplyr::left_join(elements) |>
  dplyr::mutate(OverallRating = overall$OverallRating)

outfile <- paste0(here::here("data-raw/indratings/"), indicator_name, "_ratingsumm.csv")

write.csv(ratingsumm, outfile)

flextable::flextable(ratingsumm)

```

## Comments

### Additional potential links to management in addition to uses listed above

{{MgtPotentialStart}}

[Select from list of management processes/actions using exact keywords]

{{MgtPotentialEnd}}


### What additional work would be needed for the Council to use the indicator? 

{{AdditionalWorkStart}}

[List current gaps in the indicator given the current or potentail uses listed above, what resources are needed?]

{{AdditionalWorkEnd}}


### What issues are caused if there is a gap or delay in data underlying the indicator

{{DataDelayStart}}

[Link to management process/actions and their requirements]

{{DataDelayEnd}}



