---
title:
    | 
    | SMART Indicator Report: Atlantic MultiDecadal Oscillation Index
    | 
#date: "`r format(Sys.time(), '%B %d, %Y')`"
author: ""
always_allow_html: yes
bibliography: smartind.bib
link-citations: true
site: bookdown::bookdown_site
biblio-style: apalike
github-repo: MAFMC/SMARTindicatorsData
header-includes:
  \\usepackage{array}
favicon: favicon.ico
---

```{r, include = FALSE}
## Load packages
# library(tidyverse)
library(ggplot2)
library(tidyr)
library(dplyr)
library(marea)
library(plotly)
library(DT)
library(jpeg)
library(patchwork)
library(ggthemes)
library(here)
library(zoo)
library(gt)
library(kableExtra)
library(ggiraph)

knitr::opts_chunk$set(echo = F, message = F, warning = F, fig.align = 'center', eval.after = 'fig.cap')

## Read in needed data
# Maritimes ecosystem data is here https://github.com/MarEcosystemApproaches/marea
# processed with marea_indicator_stats() and collect_marea_indicators_stats()

  mareadat <- readRDS(here::here("data-raw/mareadat.rds"))
  

## Put all mustaches here
indicator_name <- "Atlantic MultiDecadal Oscillation Index"
indicator_varname <- "ALL"

# cc_name <- "{{CC_NAME}}"
# common_name <- "{{COMMON_NAME}}"
# interactive <-  ifelse(knitr::is_latex_output(), F,T)
# 

## Pull out text and stats for indicator_name

  # check for stuff that needs escaping in indicator_name and escape it so detection works
  esc_indicator_name <- stringr::str_escape(indicator_name)
  
  mareaind <- mareadat |>
    dplyr::filter(Indicator == esc_indicator_name)
  
  
# load the data
  dataset <- unique(mareaind$Dataset)
  
  data(list = dataset, package = "marea", envir = environment())

  dat <- get(dataset)

  # dummy values to get to gdoc, will be overwritten in main body
category <- ""
specific.1 <- NA
specific.2 <- NA
specific.3 <- NA
specific.4 <- NA
specific.5 <- NA
specific.6 <- NA
measurable.1 <- NA
measurable.2 <- NA
measurable.3 <- NA
measurable.4 <- NA
measurable.5 <- NA
measurable.6 <- NA
attainable.1 <- NA
attainable.2 <- NA
attainable.3 <- NA
relevant.1 <- NA
relevant.2 <- NA
relevant.3 <- NA
timebound.1 <- NA
timebound.2 <- NA
# StartYear <- NA
# EndYear <- NA

```

# Descriptive Section

## Indicator category


Habitat-Physical
`r category <- "Habitat-Physical"`


`r paste(category)`

## Indicator name

`r paste(indicator_name)`

Includes variable(s): `r paste(indicator_varname)`

## Indicator brief description


The AMO is an observed pattern of alternating cool and warm periods of North Atlantic sea surface temperature every 40-60 years. 


`r specific.1 <- 1`

## Indicator visualization


The indicator oscillates from 1850 to 1975 and then shows a generally increasing trend.


```{r plot}

marea::plot(dat)

```


# SMART Attribute Section

## Indicator documentation

### Are indicators available for others to use (data downloadable)?

```{r download, results='asis'}
# by definition yes
if(length(dat@data$year) > 1){
  cat("Yes", "\n")
  measurable.1 <- 1
}else{
  cat("Data not in marea, need to enter info", "\n")
  measurable.1 <- 0
}
  
```

#### Where can indicators be found?

```{r where, results='asis'}

if(length(dat@data$year) > 1){
  cat("All marea indicators are here: https://github.com/MarEcosystemApproaches/marea", "\n\n")
  measurable.2 <- 1
}else{
  cat("Not in marea, need to enter info", "\n")
  measurable.2 <- 0
}
  

```

#### How often are they updated? Are future updates likely?


Updates appear to be monthly; the future is unknown


#### Who is the contact?

`r paste(maintainer("marea"))`

```{r M3}
#currently a test for existence
measurable.3 <- ifelse(is.null(maintainer("marea")), 0, 1)
```

### Gather indicator statistics

#### Units

```{r units}

specific.2 <- ifelse(length(unique(mareaind$Units)) == 0, 0, 1)
if(length(unique(mareaind$Units)) == 1 & any(is.na(mareaind$Units)|stringr::str_detect(unique(mareaind$Units), "no Units field"))) specific.2 <- 0 #either NA or no units
if(length(unique(mareaind$Units)) > 1 & any(is.na(mareaind$Units)|stringr::str_detect(unique(mareaind$Units), "no Units field"))) specific.2 <- 0.5 #some NA

# if(length(unique(clean_ecodat$Units)) == 1){
#   cat(unique(clean_ecodat$Units), "\n")
# }else{
  mareaind |>
    dplyr::select(Indicator, Units) |>
    dplyr::distinct() |>
    flextable::flextable()|>
    flextable::set_table_properties(layout = "autofit")
#}

```


#### Length of time series, start and end date, periodicity

General overview: `r paste(unique(mareaind$Time))`

```{r T1}
#currently a test for existence
timebound.1 <- ifelse(length(unique(mareaind$Time)) == 0, 0, 1)
```

Indicator specifics:

```{r}

if(any(mareaind$Varname %in% c("no Time Variables"))){
  timebound.2 <- 0 #no time variables, so not recent
  
  mareaind |>
    dplyr::select(Indicator, Varname) |>
    dplyr::distinct() |>
    flextable::flextable()|>
    flextable::set_table_properties(layout = "autofit")
  
}else{
  #end date >2023 for 2025 report = 1 Time-bound, 2020-2023 = 0.5, <2020 = 0
  # generalize to when this report is posted
  curryr <- as.numeric(format(Sys.Date(), "%Y"))
  
  maxendyr <- mareaind |>
    dplyr::select(Indicator, EPU, Varname, Value) |>
    dplyr::filter(Varname %in% c("EndYear")) |>
    tidyr::pivot_wider(names_from = Varname, values_from = Value) |>
    dplyr::select(EndYear) |>
    dplyr::distinct() |>
    dplyr::summarise(maxyr = max(EndYear, na.rm=TRUE))
    
  if(curryr-maxendyr$maxyr <= 1) timebound.2 <- 1
  if(curryr-maxendyr$maxyr >1 &  curryr-maxendyr$maxyr < 5) timebound.2 <- 0.5
  if(curryr-maxendyr$maxyr >= 5) timebound.2 <- 0

  mareaind |>
    dplyr::select(Indicator, EPU, Varname, Value) |>
    dplyr::filter(Varname %in% c("StartYear", "EndYear")) |>
    tidyr::pivot_wider(names_from = Varname, values_from = Value) |>
    #dplyr::mutate(MissingYears = EstYrs - ActualYrs) |>
    dplyr::select(Indicator, EPU, StartYear, EndYear) |> #, NumYears = ActualYrs, MissingYears) |>
    dplyr::distinct() |>
    flextable::flextable() |>
    flextable::colformat_num(big.mark = "") |>
    flextable::set_table_properties(layout = "autofit")

}

```


#### Spatial location, scale and extent

General overview: `r paste(unique(mareaind$Space))`

```{r S3}
#currently a test for existence
specific.3 <- ifelse(length(unique(mareaind$Space)) == 0, 0, 1)
```

Indicator specifics:

```{r}

  mareaind |>
    dplyr::select(Indicator, EPU) |>
    dplyr::distinct() |>
    flextable::flextable()|>
    flextable::set_table_properties(layout = "autofit")

```

#### Management scale: all species, FMP level, species level, can it be aggregated or separated to different scales?


Basinwide scale applying to the entire North Atlantic; cannot be aggregated or separated.


#### Uncertainty metrics

Uncertainty is captured in these variables:  

```{r uncert, results='asis'}

uncertvars <- "Lower95|SE|se_|SD|Upper95|Standard Error|CV|lower|upper|LCI|UCI"

unique(mareaind$Indicator[stringr::str_detect(mareaind$Indicator, uncertvars)])


#currently a test for existence
specific.4 <- ifelse(length(unique(mareaind$Indicator[stringr::str_detect(mareaind$Indicator, uncertvars)])) == 0, 0, 1)

  
```


### Are methods clearly documented to obtain source data and calculate indicators?

```{r documented, results='asis'}

if(any(stringr::str_detect(names(dat@meta), "method"))){
  cat("Yes. Methods are included in metadata", "\n")
  specific.5 <- 1
}else{
  cat("No methods documentation is provided", "\n")
  specific.5 <- 0
}

```


#### Can the indicator be calculated from current documentation?

```{r calc, results='asis'}

if(any(stringr::str_detect(names(dat@meta), "method"))){
  cat("Yes. Methods are included in metadata", "\n")
  #specific.5 <- 1
}else{
  cat("No. No methods documentation is provided", "\n")
  #specific.5 <- 0
}

```


#### Is code publicly available? up to date?

```{r codeS6, results='asis'}

# search for a github link in the metadata
if(!is.na(any(stringr::str_detect(dat@meta, "https://github.com/")))){
  cat("Yes. GitHub code repository linked", "\n")
  specific.6 <- 1
}else{
  cat("No. No link to code in metadata", "\n")
  specific.6 <- 0
}

# existence of link will be considered available, can't tell up to date

```

#### Have methods changed over time?


Unknown


### Are indicator underlying source data linked or easy to find?

```{r M4}

# search for anything in the Source field
if(!is.na(unique(mareaind$Source))){
  cat("Yes. Source information included:", unique(mareaind$Source), "\n")
  measurable.4 <- 1
}else{
  cat("No. No Source information", "\n")
  measurable.4 <- 0
}

```


#### Where are source data stored?

```{r M5}

# search for a web link in the Source field
if(!is.na(any(stringr::str_detect(mareaind$Source, "http")))){
  cat("Source data stored at", unique(mareaind$Source), "\n")
  measurable.5 <- 1
}else{
  cat("No. No link to source data", "\n")
  measurable.5 <- 0
}

```


#### How/by whom are source data updated? Are future updates likely?

```{r M6}

# search for an email in the Source field, otherwise unknown
if((any(stringr::str_detect(mareaind$Source, "@")))){
  cat("Source contact information", unique(mareaind$Source), "\n")
  measurable.6 <- 1
}else{
  cat("No. No source data steward information", "\n")
  measurable.6 <- 0
}

```


#### How often are they updated?


Unknown



## Indicator analysis/testing or history of use

### What decision or advice processes are the indicators currently used in?


None known
`r relevant.1 <- 0`


### What implications of the indicators are currently listed?


The AMO was originally described as "an internal oscillation in the climate system resulting from interactions between North Atlantic ocean currents and wind patterns." If it were an internal oscillation, it could theoretically be predicted and therefore potentially useful for forecasting. It is now thought to be "an artifact, during the historical era, of competing anthropogenic (greenhouse warming and sulphate aerosol cooling) drivers and, during the pre-historic era, an artifact of the fact that volcanic forcing happens to have displayed a roughly multidecadal pacing in past centuries." See https://michaelmann.net/the-rise-and-fall-of-the-atlantic-multidecadal-oscillation/ 
`r relevant.2 <- 0`


### Do target, limit,  or threshold values already exist for the indicator?


No.
`r relevant.3 <- 0`


### Have the indicators been tested to ensure they respond proportionally to a change in the underlying process?


Yes, the has been extensively tested using recent climate models to determine whether it is an internal climate system oscillation or not; the current conclusion is that it is not.
`r attainable.1 <- 1`


### Are the indicators sensitive to a small change in the process, or what is the threshold of change that is detectable?


Unknown
`r attainable.2 <- 0`


### Is there a time lag between the process change and the indicator change? How long?


Unknown
`r attainable.3 <- 0`


# SMART rating

```{r SMART}

indratings <- c("Described" = specific.1, "Units" = specific.2, "Spatial" = specific.3, "Uncertainty" = specific.4, 
                "Methods" = specific.5, "Code" = specific.6,
             "Available" =  measurable.1, "Online" = measurable.2, "Contact" = measurable.3, "SourceDat" = measurable.4, 
             "SourceAvail" = measurable.5, "SourceContact" = measurable.6,
             "Tested" = attainable.1, "Sensitivity" = attainable.2, "TimeLag" = attainable.3,
             "Advice" = relevant.1, "Implications" = relevant.2, "TargThresh" = relevant.3,
             "Frequency" = timebound.1, "Updated" = timebound.2)

ratings <- stack(indratings)

colnames(ratings) <- c("Rating", "Attribute")

ratings$Element <- c(rep("Specific", 6),
                     rep("Measurable", 6),
                     rep("Achievable", 3),
                     rep("Relevant", 3),
                     rep("Timebound", 2))

ratings$Indicator <- unique(indicator_name)
ratings$Category <- category

ratings <- ratings[,c("Category", "Indicator", "Element", "Attribute", "Rating")]

elements <- ratings |>
  dplyr::group_by(Element) |>
  dplyr::summarise(ElementRating = mean(Rating, na.rm=TRUE))

overall <- elements |>
  dplyr::summarise(OverallRating = mean(ElementRating, na.rm=TRUE))

ratingsumm <- ratings |>
  dplyr::left_join(elements) |>
  dplyr::mutate(OverallRating = overall$OverallRating)

outfile <- paste0(here::here("data-raw/indratings/"), indicator_name, "_ratingsumm.csv")

write.csv(ratingsumm, outfile)

flextable::flextable(ratingsumm)

```

## Comments

### Additional potential links to management in addition to uses listed above


[Select from list of management processes/actions using exact keywords]



### What additional work would be needed for the Council to use the indicator? 


[List current gaps in the indicator given the current or potentail uses listed above, what resources are needed?]



### What issues are caused if there is a gap or delay in data underlying the indicator


[Link to management process/actions and their requirements]




